---
title: "Lab_15"
author: "Lil Rascals"
date: "April 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Exams <- read.csv("StudentsPerformance.csv", sep = ",")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(cowplot)
library(stringr)
library(gapminder)
```

### Overall Question
How do the factors of gender, ethnicity, parents education, yor daily lunch, and a test prep cousre effect the affect the liklihood of a person passing the exam?
```{r}
#getting ovrall
Exam1 <- mutate(Exams, Ovr = math.score + reading.score + writing.score)
S <- length(Exam1$gender)
  #get the probability and add it to original data
#prob for race
ETPS <- Exam1 %>% filter(math.score >= 70, reading.score >= 70, writing.score >= 70) %>% group_by(race.ethnicity) %>% summarize(count = n())
ETPS <- ETPS %>% mutate(ProbE = count/S) %>% select(race.ethnicity,ProbE)
Exam1 <- Exam1 %>% left_join(ETPS, by = "race.ethnicity")
#prob for gender
GTPS <- Exam1 %>% filter(math.score >= 70, reading.score >= 70, writing.score >= 70) %>% group_by(gender) %>% summarize(count = n())
GTPS <- GTPS %>% mutate(ProbG = count/S) %>% select(gender, ProbG)
Exam1 <- Exam1 %>% left_join(GTPS, by = "gender")
#prob for parent edu
PTPS <- Exam1 %>% filter(math.score >= 70, reading.score >= 70, writing.score >= 70) %>% group_by(parental.level.of.education) %>% summarize(count = n())
PTPS <- PTPS %>% mutate(ProbP = count/S) %>% select(parental.level.of.education, ProbP)
Exam1 <- Exam1 %>% left_join(PTPS, by = "parental.level.of.education")
#prob for lunch
LTPS <- Exam1 %>% filter(math.score >= 70, reading.score >= 70, writing.score >= 70) %>% group_by(lunch) %>% summarize(count = n())
LTPS <- LTPS %>% mutate(ProbL = count/S) %>% select(lunch, ProbL)
Exam1 <- Exam1 %>% left_join(LTPS, by = "lunch")
#prob for test prep
TTPS <- Exam1 %>% filter(math.score >= 70, reading.score >= 70, writing.score >= 70) %>% group_by(test.preparation.course) %>% summarize(count = n())
TTPS <- TTPS %>% mutate(ProbT = count/S) %>% select(test.preparation.course, ProbT)
Exam1 <- Exam1 %>% left_join(TTPS, by = "test.preparation.course")
#add up probabilities
Exam1 <- Exam1 %>% mutate(PROB = (ProbE * ProbG * ProbP * ProbL * ProbT) * 10)
#graph the plot
ggplot(data = Exam1)+
  geom_smooth(aes(x = ProbG, y = Ovr, color = "Gender"), method = "lm", se = F)+
  geom_smooth(aes(x = ProbE, y = Ovr, color = "Race/Ethnicity"), method = "lm", se = F)+
  geom_smooth(aes(x = ProbP, y = Ovr, color = "Parent Education"), method = "lm", se = F)+
  geom_smooth(aes(x = ProbL, y = Ovr, color = "Lunch"), method = "lm", se = F)+
  geom_smooth(aes(x = ProbT, y = Ovr, color = "Test Prep"), method = "lm", se = F)+
  theme_bw()+
  xlab("Probability")+
  ylab("Overall Score")+
  ggtitle("The Linear Relationship of Passing Due to Factors and Thier Overall Score")+
  labs(color = "Factors")
```
###Findings:
After creating the graph and looking at the data its clear to see that the only outlier is the line represeniting the test prep catagory. The test prep catagory represents the linear relationship between the probability of a student getting a passing grade if they completed the test prep or not and their overall score. The slope is negative which would indicate that if you completed the test prep, it doesn't guarantee a passing score. After seeing this negative relationship i looked through the data and saw that there was a 50/50 on completed test preps and the students that did do it, some of them didn't pass the tests.
The other 4 factors all had positive relationships, some with a bigger coefficient than others, in which there is a larger difference of passing or not. The Lunch catagory has a big difference of probability with having two options. Therefore it can be infered about a little look through the data will show that students who have standard lunch will most liekly pass the test. The rest three, parents education, race/ethnicity, and gender show strong relationshp between having a better probability of passing and their certain traits. 
###Conclusions:
In conclusion of looking at the graph and the data there are different factors that have higher chances of passing than other ones. But for the test prep there is really no correlation between finishing it and passing the Exam.
###Recomendations:
Our recomendation is to change up the test prep course and try to make it relevant to the exam. Also try and get some help to everyone to better improve their scores so they can be on the other side of the statistic. 

### Individual SubQuestions
#### Kelsey's SubQuestion
How does being on free or reduced lunch impact your test scores?
This question is extremely important in determining whether students who are less advantaged financially do worse on their tests. This could translate to finding policies to even out student incomes and opportunities. Policies that should support these students even further should be applied if their scores are in fact lower. First I tested the scores for math, then others. 
```{r}
Exams1 <- Exams %>% filter(lunch == "free/reduced") 
perm_mean <- function(perms = 1000, n1)
{
  mean_mat <- matrix(0,1000,1)

  for (i in c(1:perms))
  {
    rsamp <- sample(Exams1$math.score, n1, replace = FALSE)
    A <- mean(rsamp)
    B <- mean(Exams$math.score)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
mean(Exams$math.score)
means <- as.tibble(perm_mean( n1=33))
sd(Exams$V1)
realdiff <- mean(Exams1$math.score)- mean(Exams$math.score)
realdiff
std <- sqrt( (sd(Exams$math.score))^2/length(Exams1$math.score))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram(binwidth=1)+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Math Test Scores with Free/Reduced Lunch")
perm_mean <- function(perms = 1000, n1)
{
  mean_mat <- matrix(0,1000,1)

  for (i in c(1:perms))
  {
    rsamp <- sample(Exams1$reading.score, n1, replace = FALSE)
    A <- mean(rsamp)
    B <- mean(Exams$reading.score)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
mean(Exams$reading.score)
means <- as.tibble(perm_mean( n1=33))
sd(Exams$V1)
realdiff <- mean(Exams1$reading.score)- mean(Exams$reading.score)
realdiff
std <- sqrt( (sd(Exams$reading.score))^2/length(Exams1$reading.score))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram(binwidth=1)+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Reading Test Scores with Free/Reduced Lunch")
Exams_math <- lm(math.score ~ reading.score, data = Exams)
Exams_math

```
From the results above it is clear that there is an enormous discrepancy between students who have free and reduced lunch and the average student's test scores. Clearly, students on free or reduced lunch have lower math scores by 7.167 points. Through a permutation test we can see that the difference is present if we run 1000 trials and that the real difference between the two is very obvious. Since our z-score is so enormous 8.9, we know that there is very little change that the difference between these two groups is due to chance alone. Therefore, we can conclude that the results are statistically significant and students with free and reduced lunch have consistently lower math test scores than other students.
I repeated this process for reading scores and found that it was still statistically significant, but that the difference in scores between the two groups was lower at 4.5 points for those students on free and reduced lunch. Our z-score was 5.8 so our data was still statistically significant. I also found out through my linear model that a higher reading score is correlated with a higher math score. Therefore, they aren't necessarily independent of eachother and higher achieving students on the math are more likely to do better on the reading tests as well. 
The new tools I used are linear model and means test. Using these tools I was able to find how test scores were impacted by free/reduced lunch. This relates to our overall question by telling us what it takes for students to do better on exams and can even help us formulate policy initiatives. 

#### James's Subquestion
```{r}
Exams3 <- Exams %>% filter(test.preparation.course == "completed") 
perm_mean2 <- function(perms = 1000, n1)
{
  mean_mat <- matrix(0,1000,1)

  for (i in c(1:perms))
  {
    rsamp1 <- sample(Exams3$math.score, n1, replace = FALSE)
    A <- mean(rsamp1)
    B <- mean(Exams$math.score)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
mean(Exams$math.score)
means <- as.tibble(perm_mean2( n1=33))
sd(Exams$V1)
realdiff <- mean(Exams3$math.score)- mean(Exams$math.score)
realdiff
std <- sqrt( (sd(Exams$math.score))^2/length(Exams3$math.score))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram(binwidth=1)+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "red", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Math Test Scores with Test Prep")

```


#### Shreeya's Subquestion
Are there differences in test scores between male and female students?

```{r}
examsGender <- Exams %>% filter(gender == "female")
examsMale <- Exams %>% filter(gender == "male")
perm_mean <- function(perms = 1000, n1)
{
  mean_mat <- matrix(0,1000,1)

  for (i in c(1:perms))
  {
    rsamp <- sample(examsGender$math.score, n1, replace = FALSE)
    A <- mean(rsamp)
    B <- mean(examsMale$math.score)
    mean_mat[[i]] <- (A-B)
  }
  return(mean_mat)
}
mean(Exams$math.score)
means <- as.tibble(perm_mean( n1=33))
sd(Exams$V1)
realdiff <- mean(examsGender$math.score)- mean(examsMale$math.score)
realdiff
std <- sqrt( (sd(examsMale$math.score))^2/length(examsGender$math.score))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram(binwidth=1)+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "orange", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Math Test Scores Between Male and Female Students")
######
perm_mean <- function(perms = 1000, n1)
{
  mean_read <- matrix(0,1000,1)

  for (i in c(1:perms))
  {
    rsamp <- sample(examsGender$reading.score, n1, replace = FALSE)
    A <- mean(rsamp)
    B <- mean(examsMale$reading.score)
    mean_read[[i]] <- (A-B)
  }
  return(mean_read)
}
mean(Exams$reading.score)
means <- as.tibble(perm_mean( n1=33))
sd(Exams$V1)
realdiff <- mean(examsGender$reading.score)- mean(examsMale$reading.score)
realdiff
std <- sqrt( (sd(examsMale$reading.score))^2/length(examsGender$reading.score))
z_score <- realdiff/std
z_score
ggplot(means, aes(x=V1 ))+geom_histogram(binwidth=1)+
  xlab("Means")+
  geom_vline(xintercept = realdiff, color = "purple", size = .3)+
  ylab("Number")+
  ggtitle("Difference in Means in Reading Test Scores Between Male and Female Students")
```
Findings:
  Using permutation tests to compare the difference in scores betweeen male and female students for the math and reading tests, I found that female students did on average five points worse than their male counterparts on the math test and about 7 points better on the reading test. The implications of these tests indicate further research into why female students are doing worse in math and better in english is necessary. A z score of 11.65 indicates these tests are statistically significant. 


###Conclusion
Overall, we found that having a student on free or reduced lunch means there is generally a lower score on the tests. We also found that on average those who completed test prep did better on the exams. 
Therefore, we would recommend implementing better anti-poverty policies in order to reduce the amount of students who rely on free and reduced lunch. Also, we could increase support for these students in ways such as having extra tutoring options for them or providing extra test prep classes for them to decrease the difference between the two groups test scores. 
In addition, we could mandate or encourage the completion of test prep courses for students since completing test prep is correlated with a better performance on tests. 

###Matthews' Subquestion:
How do test scores get affected if the student doesnt get a full meal for lunch?
```{r}
ggplot(data = Exam1)+
  geom_density(aes(x = Ovr, fill = lunch), kernel = "gaussian")+
  ggtitle("Density Curve of The Overall Score's Sorted by Their Lunches")+
  xlab("Overall Scores")+
  ylab("Density")+
  theme_bw()
```
###Findings:
Looking at the graph its clear to see that there is a bigger distribution towards students who have standard lunches compared to the free/reduced option. This can confirm that a full meal of lunch will improve your energy and will improve the test overall test scores. 


### Lab 2 Reflection


#### Kelsey's Reflection
In lab 2, one of my goals was to get better at R throughout the course of this class. I have learned so much about R and R-studio formatting and manipulating data. I have learned how to do permutation tests and gotten better at ggplot. I have come so far and learned so much throughout this class. I feel that I have gotten so much better at R and the tools it comes with. I would say my long term goals are very similar. Although I feel like now I know more about my plans for the future since I am graduating in a week. I also had my assignment for peace corps changed on me and I now leave for Lesotho in September. 

####Shreeya's Reflection
  My long term goals have not changed significantly. I think that I suceeded in my goal of learning to program in R and to use those tools to work with data more knowledgeably than I did at the start of the semester. I've learned a lot about data and what working with data means and I've come very far from when I started the class with no programming knowledge in R and very little statistical knowledge.  

###Matthew's Reflection:
In Lab 2 I stated that I wanted to learn keys skill to stand out in the workforce in which I think I have done that with learning R and getting pretty good at it. I also learned how to interpert and analyze the data to make assumptions and answer questions. My long term goals are still the same since I am not really sure of what I will try to be doing for my carrer. I came into this class after taking another computer science class and ended up hating computer science but learning R has showed me that computer science can be fun. 
